Towards the near-infrared detection of brown dwarf companions: Exploring methods to detect low mass stellar companions from blended spectra \thanks{Based on observations collected at the European Organisation for Astronomical Research in the Southern Hemisphere under ESO programme 089.C-0977(A)}

### 2018-07-17 nir_companions.tex
linting chktex warnings

% Only include extra packages if you really need them. Common packages are:
\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols
\usepackage[para]{threeparttable}
\usepackage{hyperref}
\usepackage{natbib}
\usepackage{booktabs}
\usepackage{color}
\usepackage{fix-cm}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Secondly, we mask out any remaining artifacts present in the spectra and the centers of deep telluric lines where telluric correction was not corrected properly, sometimes resulting in "emission-like" peaks in the corrected spectrum. These factors combined result in masking out around a further 10\% of the observed spectra.

@@

The binary model should provide meaningful information about the companion parameters (e.g. \(T_{\textrm{eff}}\)) and a estimate of the flux ratio of the system. These can be combined with the~\citet{baraffe_evolutionary_2003} models to constrain the mass of the companion. However, we need to be careful with this model as the inclusion of extra spectral components and associated parameters could also provide a better fit to observations which do not have a companion, by fitting components of the noise.\\

@@

We use the estimated companion temperatures from the Baraffe evolutionary models given the companion \(\rm M_2\) or \(\textrm{M}_2\sin{i}\) and stellar age as a starting point for the companion spectra grid computation and extend the grid in each direction, within the model limits. For example we show the companion temperature grid spanning \(-600\) to \(+400\)~K in Fig.~\ref{fig:Mdwarf_contours} and  \(\pm400\)~K in Figs.~\ref{fig:HD211847_simulated_contours} and \ref{fig:HD211847_result_contours}.

@@

An iterative process could be implemented to refine the RV grids, starting at a larger grid with lower RV resolution then performing a higher resolution grid about the minimum \(\chi^2\) RV values. This was only done manually but could have been automated. One could expect that a good starting RV grid step be governed by the spectral resolution, e.g. comparable to the FWHM velocity.

@@

In our examples the logg and metallicity of the synthetic models are held fixed, leaving only temperature to vary. The temperature impacts the synthetic spectral models in two main ways: the flux level of the continuum; and the number and strength of the absorption lines.  In the binary model the contributions from the individual components is scaled by the flux ratio. If the temperature of the companion increases then the flux and radius of the companion increases. Its contribution of the companion to the binary model increases and the flux ratio \(\rm F_1/F_2\) decreases. This effectively makes the lines of the spectrum of the host component relatively smaller in the normalized binary model spectrum. Due to the initial mismatch of synthetic spectral lines, a decrease in relative strength of the host lines decreases the \(\chi^2\) value, and is a better "match" to the observation. This causes the recovered temperature of the companion to be higher than expected, >2000~K higher if allowed by the exploration grid. The \(\chi^2\) approach is dominated by reducing the mismatch in the spectrum of the host rather than detecting the spectra of the companion.
This spectral mismatch is not observed in the simulations in Sect.~\ref{subsec:simulated_binaries} since they are created using the synthetic spectra themselves, and hence they do recover the correct host spectra.

@@

In Table~\ref{tab:line_contributions} we calculate some properties of the spectral lines in the PHOENIX-ACES library between 2110--2160~nm. We count the number of spectral lines (\emph{no. lines}) deeper than 5\%, and take the average depth (\emph{avg.\ depth}) of these lines. The contribution \emph{cont. depth} of the companion lines to a combined spectrum accounts for the flux ratio between the two components. Here we use a Sun-like host with \(T_{\textrm{eff}_1}=5800\)~K. This simplified combination neglects the continuum shapes of both spectral components and uses the average flux ratios for this wavelength range. The PHOENIX-ACES spectra in the temperature range of 2500--5000~K shown in Fig.~\ref{fig:comp_spectra} can be used to get a visual indication of the line density and depth measured here.

@@

The bump observed at 5100~K in the \(\chi^2\) curves is due to a discontinuity in the PHOENIX-ACES modeling. The "reference wavelength defining the mean optical depth grid" is changed at 5000~K~\citep[][Sect. 2.3]{husser_new_2013}. Care needs to be taken if trying to detect a companion near this temperature.

@@

\subsubsection{Component RV separation}
\label{subsubsec:rv_seperation}
Another factor which could contribute to an unsuccessful detection is the RV separation between the host and companion, \(rv_2\). Estimates for our observations are given in the last column of Table~\ref{tab:observations}. If \({rv}_2\) is small compared to the line width, then all the same lines of both components will be blended. This is indeed the case for {HD 4747}, {HD 211847}, and {HD 202206} with expected \(|{rv}_2| < 2\)~kms\(^{-1}\). This may have contributed to the lack of recovery with both components of the binary model trying to fit to the same features. This may even cause some correlation between the parameters of the two components. The RV separation of the two components changes with orbital phase. Having multiple spectra of the same target distributed in phase may allow the RV of the spectral components to be better recovered~\citep [e.g. ][]{czekala_disentangling_2017, sablowski_spectral_2016}.

@@

\subsubsection{The BT-Settl models}
\label{subsubsec:bt-settl}
We note that the PHOENIX-ACES models are not the only spectral libraries available with the other notable library considered for this work is the BT-Settl models, ~\citep{allard_model_2010,allard_btsettl_2013,baraffe_new_2015}. The included modeling of dust and cloud formation, as well as hydro-dynamical modeling atmospheric mixing/settling for atmospheres with \(T_{\textrm{eff}}\) below \(\sim2600\)~K, make the BT-Settl models valid across the regime from stars to BDs as cool as 400~K. As the BT-Settl models are suitable to model the atmospheres of the brown dwarfs they would be useful for the companion recovery technique developed here. However, as shown in Sect.~\ref{subsection:results-hd211847} and \ref{subsection:injection-recovery}, we were unable to successfully recover the 155 \(\rm M_J\) (\(T_{\textrm{eff}}\sim3200\)~K) low mass star companion of {HD 211847} and derived a temperature upper limit for our methodology of around 3800~K. These are both well above the 2300~K cutoff of the PHOENIX-ACES models and for the onset of dust- and cloud-formation phenomena, at 2600 K..

@@

On the modeling side, there are continual improvements in atmospheric modeling and their associated synthetic spectral libraries: as seen with the evolution of the BT-Settl models~\cite{allard_btsettl_2013}. With additional physics and improved line lists and solar abundances~\citep [e.g.][]{asplund_chemical_2009,caffau_solar_2011}, the synthetic libraries are reaching a better agreement with nIR observations. An improved agreement between the nIR observations and synthetic spectra will be crucial to improve the performance of the spectral recovery technique presented here.

@@

The long orbital periods of these targets is also a contributing factor to the insufficient separations. Most of the targets observed here have orbital periods much longer than an observing semester (183 days). An optimal pair of observations (achieved at the extrema) would need to have been obtained from separate observing periods (between 2 months and 19 years apart). In some cases, even observations taken at the beginning and end of the semester would not be sufficient to achieve companion separation (depending on the phase). Requiring separate observing periods to even achieve the minimum \(\Delta \rm RV\) larger than the line FWHM.  At the time it was impossible to ask for time over several semesters in a regular proposal.

@@

For future observations we suggest that the known orbital solution of the companion be used to estimate the companions' RV curve during the observing period, with the companion \(M_2\sin{i}\) providing an RV upper-limit. Knowing the instrumental wavelength and resolution, a constraint can then be set to avoid taking observations when the companion spectra are insufficiently separated, or \(\Delta RV\) < FWHM. This constraint can be set using the absolute and relative \emph{time-link} constraints available in ESO's Phase 2 Proposal Preparation (P2PP) tool.
Additionally, analyzing the known orbital solution before-hand, to determine RV constraints will also help identify the best time to observe, if observations from separate periods will be required or, if an optimally separated companion differential is even feasible.

@@

% The best way to enter references is to use BibTeX:
\bibliographystyle{mnras}
\bibliography{nir_paper}

@@

\begin{appendix} % First appendix

@@

    It is necessary to have a consistent instrumental setup~\citet{ferluga_separating_1997}, to avoid introducing extra instrumental effects (e.g. slit-width and/or filters) into the spectral differentials and to always observe the same wavelength range and maximize the information to be extracted. In our case, the second observation of {HD 202206} and fourth of {HD 30501} were taken with different filters compared to the other observations. Therefore, these two observations could not be used for this differential analysis. As noted in~\citep{hadrava_disentangling_2009}, any spectral differences in the filters would add extra unknown signal/noise making it harder to disentangle the faint spectral differences.

@@

    The error on estimated RV values is calculated using the general error propagation formula~\citep{ku_notes_1966} with the RV equation and the errors on the orbital parameters. For a function, \(f\), with errors on the inputs \(\delta x\), \(\delta y\) etc, it follows: 
    \begin{align}
    f &= f(x, y, z, ...)\\
    \delta f &= \sqrt{{\left( \frac{\partial f}{\partial x} \delta x\right)}^2 +  {\left(\frac{\partial f}{\partial y} \delta y\right)}^2 + {\left(\frac{\partial f}{\partial z} \delta z\right)}^2 + ...}
    \end{align}

    \section{Estimating Companion-host Flux ratio}
    \label{compaion flux ration}
    The companion-host flux or contrast ratio of the systems are calculated using \( \frac{F_{2}}{F_{1}} \approx 2.512^{m_{1}-m_{2}} \), where \(m_{1} \) and \(m_{2} \) are the magnitude of the host and companion respectively.\ \(m_{1} \) is the magnitude of the host in the literature, while \(m_{2} \) is obtained from stellar evolution models of~\citet{baraffe_evolutionary_2003, baraffe_new_2015}, using the companion mass (or \(M_{2}\sin{i}\))) and the host's age. If only the companion minimum mass is known then this will correspond to the flux ratio lower limit, (or worst case scenario).  The models, interpolated to the companion mass, also provide a first estimate of companion properties such as \(T_{\textrm{eff}}\), logg, \(R/R_{\sun}\) and the magnitude in many wavelength bands.
    We use the \textit{K}-band magnitude specifically for this work to estimate the flux ratio between the host and companion using the host \textit{K}-band magnitude obtained from SIMBAD~\citep{wenger_simbad_2000}. Our flux ratio estimates are presented in table~\ref{tab:flux_table}.
    The companion \(T_{\textrm{eff}}\) and logg values specifically are used to influence the selection of synthetic model grids to perform the \(\chi^2\) analysis over.

@@

\end{appendix}

@@

% Don't change these lines
\bsp	% typesetting comment
\label{lastpage}
\end{document}

>>>>

% Only include extra packages if you really need them. Common packages are:
\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
%\usepackage{amssymb}	% Extra maths symbols
\usepackage[para]{threeparttable}
\usepackage{hyperref}
\usepackage{natbib}
\usepackage{booktabs}
\usepackage{color}
\usepackage{fix-cm}
%\usepackage{caption}   % Caption is required to aviod errors.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Secondly, we mask out any remaining artifacts present in the spectra and the centers of deep telluric lines where telluric correction was not corrected properly, sometimes resulting in ``emission-like'' peaks in the corrected spectrum. These factors combined result in masking out around a further 10\% of the observed spectra.

@@

The binary model should provide meaningful information about the companion parameters (e.g.\ \(T_{\textrm{eff}}\)) and a estimate of the flux ratio of the system. These can be combined with the~\citet{baraffe_evolutionary_2003} models to constrain the mass of the companion. However, we need to be careful with this model as the inclusion of extra spectral components and associated parameters could also provide a better fit to observations which do not have a companion, by fitting components of the noise.\\

@@

We use the estimated companion temperatures from the Baraffe evolutionary models given the companion \(\rm M_2\) or \(\textrm{M}_2\sin{i}\) and stellar age as a starting point for the companion spectra grid computation and extend the grid in each direction, within the model limits. For example we show the companion temperature grid spanning \(-600\) to \(+400\)~K in Fig.~\ref{fig:Mdwarf_contours} and \(\pm400\)~K in Figs.~\ref{fig:HD211847_simulated_contours} and~\ref{fig:HD211847_result_contours}.

@@

An iterative process could be implemented to refine the RV grids, starting at a larger grid with lower RV resolution then performing a higher resolution grid about the minimum \(\chi^2\) RV values. This was only done manually but could have been automated. One could expect that a good starting RV grid step be governed by the spectral resolution, e.g.\ comparable to the FWHM velocity.

@@

In our examples the logg and metallicity of the synthetic models are held fixed, leaving only temperature to vary. The temperature impacts the synthetic spectral models in two main ways: the flux level of the continuum; and the number and strength of the absorption lines. In the binary model the contributions from the individual components is scaled by the flux ratio. If the temperature of the companion increases then the flux and radius of the companion increases. Its contribution of the companion to the binary model increases and the flux ratio \(\rm F_1/F_2\) decreases. This effectively makes the lines of the spectrum of the host component relatively smaller in the normalized binary model spectrum. Due to the initial mismatch of synthetic spectral lines, a decrease in relative strength of the host lines decreases the \(\chi^2\) value, and is a better ``match'' to the observation. This causes the recovered temperature of the companion to be higher than expected, >2000~K higher if allowed by the exploration grid. The \(\chi^2\) approach is dominated by reducing the mismatch in the spectrum of the host rather than detecting the spectra of the companion.
This spectral mismatch is not observed in the simulations in Sect.~\ref{subsec:simulated_binaries} since they are created using the synthetic spectra themselves, and hence they do recover the correct host spectra.

@@

In Table~\ref{tab:line_contributions} we calculate some properties of the spectral lines in the PHOENIX-ACES library between 2110--2160~nm. We count the number of spectral lines (\emph{no.\ lineses}) deeper than 5\%, and take the average depth (\emph{avg.\ depth}) of these lines. The contribution \emph{cont.\ depth} of the companion lines to a combined spectrum accounts for the flux ratio between the two components. Here we use a Sun-like host with \(T_{\textrm{eff}_1}=5800\)~K. This simplified combination neglects the continuum shapes of both spectral components and uses the average flux ratios for this wavelength range. The PHOENIX-ACES spectra in the temperature range of 2500--5000~K shown in Fig.~\ref{fig:comp_spectra} can be used to get a visual indication of the line density and depth measured here.

@@

The bump observed at 5100~K in the \(\chi^2\) curves is due to a discontinuity in the PHOENIX-ACES modeling. The ``reference wavelength defining the mean optical depth grid'' is changed at 5000~K~\citep[][Sect. 2.3]{husser_new_2013}. Care needs to be taken if trying to detect a companion near this temperature.

@@

\subsubsection{Component RV separation}
\label{subsubsec:rv_seperation}
Another factor which could contribute to an unsuccessful detection is the RV separation between the host and companion, \(rv_2\). Estimates for our observations are given in the last column of Table~\ref{tab:observations}. If \({rv}_2\) is small compared to the line width, then all the same lines of both components will be blended. This is indeed the case for {HD 4747}, {HD 211847}, and {HD 202206} with expected \(|{rv}_2| < 2\)~kms\(^{-1}\). This may have contributed to the lack of recovery with both components of the binary model trying to fit to the same features. This may even cause some correlation between the parameters of the two components. The RV separation of the two components changes with orbital phase. Having multiple spectra of the same target distributed in phase may allow the RV of the spectral components to be better recovered~\citep[e.g.][]{czekala_disentangling_2017, sablowski_spectral_2016}.

@@

\subsubsection{The BT-Settl models}
\label{subsubsec:bt-settl}
We note that the PHOENIX-ACES models are not the only spectral libraries available with the other notable library considered for this work is the BT-Settl models, ~\citep{allard_model_2010,allard_btsettl_2013,baraffe_new_2015}. The included modeling of dust and cloud formation, as well as hydro-dynamical modeling atmospheric mixing/settling for atmospheres with \(T_{\textrm{eff}}\) below \(\sim2600\)~K, make the BT-Settl models valid across the regime from stars to BDs as cool as 400~K. As the BT-Settl models are suitable to model the atmospheres of the brown dwarfs they would be useful for the companion recovery technique developed here. However, as shown in Sect.~\ref{subsection:results-hd211847} and~\ref{subsection:injection-recovery}, we were unable to successfully recover the 155 \(\rm M_J\) (\(T_{\textrm{eff}}\sim3200\)~K) low mass star companion of {HD 211847} and derived a temperature upper limit for our methodology of around 3800~K. These are both well above the 2300~K cutoff of the PHOENIX-ACES models and for the onset of dust- and cloud-formation phenomena, at 2600 K..

@@

On the modeling side, there are continual improvements in atmospheric modeling and their associated synthetic spectral libraries: as seen with the evolution of the BT-Settl models~\cite{allard_btsettl_2013}. With additional physics and improved line lists and solar abundances~\citep[e.g.][]{asplund_chemical_2009,caffau_solar_2011}, the synthetic libraries are reaching a better agreement with nIR observations. An improved agreement between the nIR observations and synthetic spectra will be crucial to improve the performance of the spectral recovery technique presented here.

@@

The long orbital periods of these targets is also a contributing factor to the insufficient separations. Most of the targets observed here have orbital periods much longer than an observing semester (183 days). An optimal pair of observations (achieved at the extrema) would need to have been obtained from separate observing periods (between 2 months and 19 years apart). In some cases, even observations taken at the beginning and end of the semester would not be sufficient to achieve companion separation (depending on the phase). Requiring separate observing periods to even achieve the minimum \(\Delta \rm RV\) larger than the line FHWM.\@At the time it was impossible to ask for time over several semesters in a regular proposal.

@@

For future observations we suggest that the known orbital solution of the companion be used to estimate the companions' RV curve during the observing period, with the companion \(M_2\sin{i}\) providing an RV upper-limit. Knowing the instrumental wavelength and resolution, a constraint can then be set to avoid taking observations when the companion spectra are insufficiently separated, or \(\Delta RV\) < FHWM.\@This constraint can be set using the absolute and relative \emph{time-link} constraints available in ESO's Phase 2 Proposal Preparation (P2PP) tool.
Additionally, analyzing the known orbital solution before-hand, to determine RV constraints will also help identify the best time to observe, if observations from separate periods will be required or, if an optimally separated companion differential is even feasible.

@@

% The best way to enter references is to use BibTeX:
\bibliographystyle{mnras}
\bibliography{nir_companions}

@@

@@

    It is necessary to have a consistent instrumental setup~\citet{ferluga_separating_1997}, to avoid introducing extra instrumental effects (e.g.\ slit-width and/or filters) into the spectral differentials and to always observe the same wavelength range and maximize the information to be extracted. In our case, the second observation of {HD 202206} and fourth of {HD 30501} were taken with different filters compared to the other observations. Therefore, these two observations could not be used for this differential analysis. As noted in~\citep{hadrava_disentangling_2009}, any spectral differences in the filters would add extra unknown signal/noise making it harder to disentangle the faint spectral differences.

@@

    The error on estimated RV values is calculated using the general error propagation formula~\citep{ku_notes_1966} with the RV equation and the errors on the orbital parameters. For a function, \(f\), with errors on the inputs \(\delta x\), \(\delta y\) etc, it follows: 
    \begin{align}
    f &= f(x, y, z, \ldots)\\
    \delta f &= \sqrt{{\left( \frac{\partial f}{\partial x} \delta x\right)}^2 + {\left(\frac{\partial f}{\partial y} \delta y\right)}^2 + {\left(\frac{\partial f}{\partial z} \delta z\right)}^2 + \ldots}
    \end{align}

    \section{Estimating Companion-host Flux ratio}
    \label{compaion flux ration}
    The companion-host flux or contrast ratio of the systems are calculated using \( \frac{F_{2}}{F_{1}} \approx 2.512^{m_{1}-m_{2}} \), where \(m_{1} \) and \(m_{2} \) are the magnitude of the host and companion respectively.\ \(m_{1} \) is the magnitude of the host in the literature, while \(m_{2} \) is obtained from stellar evolution models of~\citet{baraffe_evolutionary_2003, baraffe_new_2015}, using the companion mass (or \(M_{2}\sin{i}\)) and the host's age. If only the companion minimum mass is known then this will correspond to the flux ratio lower limit, (or worst case scenario). The models, interpolated to the companion mass, also provide a first estimate of companion properties such as \(T_{\textrm{eff}}\), logg, \(R/R_{\sun}\) and the magnitude in many wavelength bands.
    We use the \textit{K}-band magnitude specifically for this work to estimate the flux ratio between the host and companion using the host \textit{K}-band magnitude obtained from SIMBAD~\citep{wenger_simbad_2000}. Our flux ratio estimates are presented in table~\ref{tab:flux_table}.
    The companion \(T_{\textrm{eff}}\) and logg values specifically are used to influence the selection of synthetic model grids to perform the \(\chi^2\) analysis over.

@@

@@

% Don't change these lines
\bsp{}	% typesetting comment
\label{lastpage}
\end{document}
